### 2020-08-25 tables.Rmd
started analysis dxa and us

---
title: "tables"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

source("./R/libs.R")

```

```{r table1-characteristics}

# Read dxa data

dxa_whole <- read_excel("./data/tr010_dxa_standarddefinitions.xlsx") %>%
  dplyr::select(Etternavn, Kjønn,Målingsdato,Alder, Vekt:`HeleVev%fett`) %>%
  pivot_longer(names_to = "variable", values_to = "value", cols = Vekt:`HeleVev%fett`) %>%
  mutate(value = gsub("kg", "", value), 
         value = gsub("[[:space:]]", "", value, fixed = FALSE), 
         value = as.numeric(value)) %>%

  separate(Etternavn, into = c("Study", "participant")) %>%
  dplyr::select(participant, Date = Målingsdato, variable, value) %>%
  filter(!(participant %in% c("P08", "P20", "P17"))) %>%
  group_by(participant) %>%
  mutate(time = if_else(Date == min(Date), "S0", 
                        if_else(Date == max(Date), "post1w", 
                                "S12")), 
         leg = if_else(variable == "BeinhøyreMagermasse", "R", "L")) %>%
  print()

# DXA weight stature

read_excel("./data/tr010_dxa.xlsx") %>%
    inner_join(read_excel("./data/leg_randomization.xlsx") %>%
               dplyr::select(participant, sex)) %>%

  group_by(participant, sex, group) %>%
  summarise(mass = mean(bodymass_manual.weight, na.rm = TRUE), 
            stature = mean(height, na.rm = TRUE)) %>%
  group_by(group, sex) %>%
  summarise(mass = mean(stature), 
            stature = mean(stature), 
            n = n()) %>%

  print()

```